---
title: "sjt2164 Homework 1"
output: github_document
---

## Question 1

_This question uses the `tidyverse` library and the `Penguins` dataset. The codes for loading the library and the dataset are hidden using `include = FALSE`._

```{r setup, include = FALSE}
library(tidyverse)
```

```{r data, include = FALSE}
data("penguins", package = "palmerpenguins")
```

```{r summary, include = FALSE}
summary(penguins) #overview of dataset.Tells us how many of each variable contains "NA".
```

**Summary of Dataset**

The `Penguins` dataset (`r ncol(penguins)` columns x `r nrow(penguins)` rows), contains data on `r nrow(penguins)` penguins with the following variables:

* **Factor variables** include `species` (`r levels(pull(penguins,species))`), `island` (`r levels(pull(penguins,island))`), `year` (`r min(pull(penguins,year))` - `r max(pull(penguins,year))`), and `sex` (`r levels(pull(penguins,sex))`). 
* **Numeric variables** include doubles (`bill_length_mm` and `bill_depth_mm`) and integers (`body_mass_g` and `flipper_length_mm`). 

Using the `summary(penguins)` code, we also see that there are several "NA" values in the dataset. Each of the numberic variables listed above had 2 NA values. 11 of the 344 penguins had NA values for `sex`.

Body mass is measured in grams and ranges from `r min(pull(penguins,body_mass_g), na.rm=TRUE)`g to `r max(pull(penguins,body_mass_g), na.rm=TRUE)`g with a mean of `r round(mean(pull(penguins,body_mass_g),na.rm= TRUE), digits=2)`g. Bill length (min: `r min(pull(penguins,bill_length_mm), na.rm=TRUE)`, max: `r max(pull(penguins,bill_length_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,bill_length_mm),na.rm= TRUE), digits=2)`), bill depth (min: `r min(pull(penguins,bill_depth_mm), na.rm=TRUE)`, max: `r max(pull(penguins,bill_depth_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,bill_depth_mm),na.rm= TRUE), digits=2)`), and flipper length (min: `r min(pull(penguins,flipper_length_mm), na.rm=TRUE)`, max: `r max(pull(penguins,flipper_length_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,flipper_length_mm),na.rm= TRUE), digits=2)`) are all measured in millimeters.

**Penguins Scatterplot**

_Note: The code, messages, and any warning notices are hidden using `echo = FALSE`, `message = FALSE`, and `warning = FALSE`. Only the scatterplot is shown below._

Using the `Penguins` dataset, a scatterplot is generated below using `ggplot() + geom_point()`. The distribution of penguin species is shown bellow with `flipper_length_mm` as the y-variable and `bill_length_mm` as the x-variable. Each penguin species is represented by a different color as depicted in the key of the scatterplot.

Since 2 penguins had "NA" values for both `flipper_length_mm` and `bill_length_mm`, they are excluded. Thus, the scatterplot contains data for a total of 342 penguins. 

The scatterplot is exported using `ggsave`.

**Scatterplot (n = 342)**

```{r scatterplot, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
#There is a warning message: "removed 2 rows containing missing values or values outside the scale range (`geom_point()`).

ggsave("penguins_scatterplot.pdf")
```

## Question 2

The code below shows how the data frame `question2_df` is constructed. This data frame consists of the following variables:

* `sample_value` = a random sample of size 10 from a standard Normal distribution
* `value_greater_than_0` = a logical vector that indicates whether elements of `sample_value` are greater than 0
* `nycity_names` = a character vector of length 10
* `top_industries` = a factor vector of length 10 containing 3 levels (`entertainment`, `healthcare`, and `technology`). 

```{r dataframe setup}
question2_df = 
    tibble(
      sample_value = rnorm(10),
      value_greater_than_0 = ifelse(sample_value > 0, "Yes", "No"),
      nycity_names = c("Nyc", "Buffalo", "Rochester", "Yonkers", "Syracuse", "Albany", "New Rochelle", "Schenectady", "Utica", "Mount_Vernon"),
      top_industries = factor(c("entertainment", "healthcare", "technology", "healthcare", "healthcare", "healthcare", "technology", "technology", "technology", "healthcare"))
    )
```

**Taking the Mean**

Using the `pull` function to take the mean of the variables in the `question2_df` data frame, we get:

* `sample_value` = `r mean(pull(question2_df,sample_value))`
* `value_greater_than_0` = `r mean(pull(question2_df,value_greater_than_0))`
* `nycity_names` = `r mean(pull(question2_df,nycity_names))`
* `top_industries` = `r mean(pull(question2_df,top_industries))`

As shown above, the mean of `sample_value` can be calculated. However, the mean cannot be calculated for the remaining 3 variables. 

_It is also important to note that, in this case, a mean cannot be calculated for the logical vector because "Yes" and "No" is used in the set up. Had "Yes" been "TRUE" and "No" been "FALSE", then it is possible to calculate the mean for `value_greater_than_0`. This is because in Rstudio, TRUE = 1 and FALSE is 0._

**Converting Variable Types**

To convert the logical, character, and factor variables, we can use the function `as.numeric`. The output is hidden using `eval = FALSE`.

```{r converting variables, eval = FALSE}

q2_df_updated = 
  question2_df|>
    mutate(
      over_zero = as.numeric(value_greater_than_0),
      names_updated = as.numeric(nycity_names),
      indus_updated = as.numeric(top_industries)
    )
```


* `sample_value` = `r mean(pull(question2_df,sample_value))`
* `value_greater_than_0` = `r mean(pull(question2_df,value_greater_than_0))`
* `nycity_names` = `r mean(pull(question2_df,nycity_names))`
* `top_industries` = `r mean(pull(question2_df,top_industries))`


What happens, and why? Does this help explain what happens when you try to take the mean?

