---
title: "sjt2164 Homework 1"
output: github_document
---

### Question 1

Question 1 uses the `tidyverse` library and the `Penguins` dataset. The codes for loading the library and the dataset are hidden using `include = FALSE` for easier reading.

```{r setup, include = FALSE}
library(tidyverse)
```

```{r data, include = FALSE}
data("penguins", package = "palmerpenguins")
```

```{r, include = FALSE}
summary(penguins) #overview of dataset.Tells us how many of each variable contains "NA".
```

The `Penguins` dataset (`r ncol(penguins)` columns x `r nrow(penguins)` rows), contains data on `r nrow(penguins)` penguins. Data collected on these penguins include:

* `species`(factor variable): `r levels(pull(penguins,species))` 
* `island`(factor variable): `r levels(pull(penguins,island))`
* `bill_length_mm`(double variable) in milimeteres
* `bill_depth_mm` (double variable) in milimeteres
* `flipper_length_mm` (integer variable) in milimeteres
* `body_mass_g` (integer variable) in grams
* `sex` (factor variable): `r levels(pull(penguins,sex))`
* `year` (factor variable): `r min(pull(penguins,year))` - `r max(pull(penguins,year))`

The mean flipper length is `r round(mean(pull(penguins,flipper_length_mm),na.rm= TRUE), digits=2)` mm.
