---
title: "P8105 Homework 1"
author: Serena Ting (sjt2164) 
date: 9/21/2024
output: github_document
---

## Question 1

_This question uses the `tidyverse` library and the `Penguins` dataset. The codes for loading the library and the dataset are hidden using `include = FALSE`._

```{r setup, include = FALSE}
library(tidyverse)
```

```{r data, include = FALSE}
data("penguins", package = "palmerpenguins")
```

```{r summary, include = FALSE}
summary(penguins) #overview of dataset.Tells us how many of each variable contains "NA".
```

**Summary of Dataset**

The `Penguins` dataset (`r ncol(penguins)` columns x `r nrow(penguins)` rows), contains data on `r nrow(penguins)` penguins with the following variables:

* **Factor variables** include `species` (`r levels(pull(penguins,species))`), `island` (`r levels(pull(penguins,island))`), `year` (`r min(pull(penguins,year))` - `r max(pull(penguins,year))`), and `sex` (`r levels(pull(penguins,sex))`). 
* **Numeric variables** include doubles (`bill_length_mm` and `bill_depth_mm`) and integers (`body_mass_g` and `flipper_length_mm`). 

Using the `summary(penguins)` code, we also see that there are several "NA" values in the dataset. Each of the numberic variables listed above had 2 NA values. 11 of the 344 penguins had NA values for `sex`.

Body mass is measured in grams and ranges from `r min(pull(penguins,body_mass_g), na.rm=TRUE)`g to `r max(pull(penguins,body_mass_g), na.rm=TRUE)`g with a mean of `r round(mean(pull(penguins,body_mass_g),na.rm= TRUE), digits=2)`g. Bill length (min: `r min(pull(penguins,bill_length_mm), na.rm=TRUE)`, max: `r max(pull(penguins,bill_length_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,bill_length_mm),na.rm= TRUE), digits=2)`), bill depth (min: `r min(pull(penguins,bill_depth_mm), na.rm=TRUE)`, max: `r max(pull(penguins,bill_depth_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,bill_depth_mm),na.rm= TRUE), digits=2)`), and flipper length (min: `r min(pull(penguins,flipper_length_mm), na.rm=TRUE)`, max: `r max(pull(penguins,flipper_length_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,flipper_length_mm),na.rm= TRUE), digits=2)`) are all measured in millimeters.

**Penguins Scatterplot**

_Note: The code, messages, and any warning notices are hidden using `echo = FALSE`, `message = FALSE`, and `warning = FALSE`. Only the scatterplot is shown below._

Using the `Penguins` dataset, a scatterplot is generated below using `ggplot() + geom_point()`. The distribution of penguin species is shown bellow with `flipper_length_mm` as the y-variable and `bill_length_mm` as the x-variable. Each penguin species is represented by a different color as depicted in the key of the scatterplot.

Since 2 penguins had "NA" values for both `flipper_length_mm` and `bill_length_mm`, they are excluded. Thus, the scatterplot contains data for a total of 342 penguins. 

The scatterplot is exported using `ggsave`.

**Scatterplot (n = 342)**

```{r scatterplot, echo = FALSE, message = FALSE, warning = FALSE}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
#There is a warning message: "removed 2 rows containing missing values or values outside the scale range (`geom_point()`).

ggsave("penguins_scatterplot.pdf")
```

## Question 2

The code below shows how the data frame `question2_df` is constructed. This data frame consists of the following variables:

* `sample_value` (_double_) = a random sample of size 10 from a standard Normal distribution
* `value_greater_than_0` (_character_) = a logical vector that indicates whether elements of `sample_value` are greater than 0
* `nycity_names` (_character_) = a character vector of length 10
* `top_industries` (_factor_) = a factor vector of length 10 containing 3 levels (`entertainment`, `healthcare`, and `technology`). 

```{r dataframe setup}
question2_df = 
    tibble(
      sample_value = rnorm(10),
      value_greater_than_0 = ifelse(sample_value > 0, TRUE, FALSE),
      nycity_names = c("Nyc", "Buffalo", "Rochester", "Yonkers", "Syracuse", "Albany", "New Rochelle", "Schenectady", "Utica", "Mount_Vernon"),
      top_industries = factor(c("entertainment", "healthcare", "technology", "healthcare", "healthcare", "healthcare", "technology", "technology", "technology", "healthcare"))
    )
```

**Taking the Mean**

Using the `pull` function to take the mean of the variables in the `question2_df` data frame, we get:

* `sample_value` = `r mean(pull(question2_df,sample_value))`
* `value_greater_than_0` = `r mean(pull(question2_df,value_greater_than_0))`
* `nycity_names` = `r mean(pull(question2_df,nycity_names))`
* `top_industries` = `r mean(pull(question2_df,top_industries))`

As shown above, the mean of `sample_value` and `value_greater_than_0` can be calculated. `sample_value` is a double and thus, mean can be calcuated. We can also calculate the mean of `value_greater_than_0` because TRUE = 1 and FALSE = 0 in logical vectors. However, the mean cannot be calculated for the remaining 2 variables: `nycity_names` (character) and `top_industries` (factor). 

**Converting Variable Types**

The code below shows an attempt to convert the logical, character, and factor variables using the function `as.numeric`. The output of the codes below are hidden using `eval = FALSE`.

```{r converting variables, eval = FALSE}
q2_df_updated = 
  question2_df|>
    mutate(
      over_zero = as.numeric(value_greater_than_0),
      names_updated = as.numeric(nycity_names),
      indus_updated = as.numeric(top_industries)
    )
```

After running the code above as well as `view(q2_df_updated)` in the console, we see that values for `names_updated` are all "NA". This is reflected in the warning messages in `mutate()`. The character vector cannot be converted using `as.numeric` because of "NAs introduced by coercion". On the other hand, the values for `indus_updated` and `over_zero` were both converted into numbers. 

**Mean Calculations after Conversion Attempt**

```{r converted mean, eval = FALSE}
mean(pull(q2_df_updated,over_zero))
mean(pull(q2_df_updated,names_updated))
mean(pull(q2_df_updated,indus_updated))
```

Running the above code chunk, we find that the mean of `indus_updated` and `over_zero` can be calculated. Since NAs were introduced for `names_updated`, this explains why we are still unable to calculate the mean after attempting to convert the variable with the `as.numeric` function.
