---
title: "sjt2164 Homework 1"
output: github_document
---

### Question 1

_This question uses the `tidyverse` library and the `Penguins` dataset. The codes for loading the library and the dataset are hidden using `include = FALSE`._

```{r setup, include = FALSE}
library(tidyverse)
```

```{r data, include = FALSE}
data("penguins", package = "palmerpenguins")
```

```{r, include = FALSE}
summary(penguins) #overview of dataset.Tells us how many of each variable contains "NA".
```

**Summary of Dataset**

The `Penguins` dataset (`r ncol(penguins)` columns x `r nrow(penguins)` rows), contains data on `r nrow(penguins)` penguins with the following variables:

* **Factor variables** include `species` (`r levels(pull(penguins,species))`), `island` (`r levels(pull(penguins,island))`), `year` (`r min(pull(penguins,year))` - `r max(pull(penguins,year))`), and `sex` (`r levels(pull(penguins,sex))`). 
* **Numeric variables** include doubles (`bill_length_mm` and `bill_depth_mm`) and integers (`body_mass_g` and `flipper_length_mm`). 

Using the `summary(penguins)` code, we also see that there are several "NA" values in the dataset. Each of the numberic variables listed above had 2 NA values. 11 of the 344 penguins had NA values for `sex`.

Body mass is measured in grams and ranges from `r min(pull(penguins,body_mass_g), na.rm=TRUE)`g to `r max(pull(penguins,body_mass_g), na.rm=TRUE)`g with a mean of `r round(mean(pull(penguins,body_mass_g),na.rm= TRUE), digits=2)`g. Bill length (min: `r min(pull(penguins,bill_length_mm), na.rm=TRUE)`, max: `r max(pull(penguins,bill_length_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,bill_length_mm),na.rm= TRUE), digits=2)`), bill depth (min: `r min(pull(penguins,bill_depth_mm), na.rm=TRUE)`, max: `r max(pull(penguins,bill_depth_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,bill_depth_mm),na.rm= TRUE), digits=2)`), and flipper length (min: `r min(pull(penguins,flipper_length_mm), na.rm=TRUE)`, max: `r max(pull(penguins,flipper_length_mm), na.rm=TRUE)`, mean: `r round(mean(pull(penguins,flipper_length_mm),na.rm= TRUE), digits=2)`) are all measured in millimeters.

## Penguins Scatterplot

_Using the `Penguins` dataset, the code below generates a scatterplot with `flipper_length_mm` as the y-variable and `bill_length_mm` as the x-variable. Each penguin species is represented by a different color as depicted in the key of the scatterplot._

```{r plot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point()
#There is a warning message: "removed 2 rows containing missing values or values outside the scale range (`geom_point()`).
```
